Joint Neighborhood Subgraphs Link Prediction

### 2017-08-01 paper.tex
abstract and corrections

\begin{abstract}
The amount of relational data is increasing rapidly. These datasets are best represented as networks or graphs. A key characteristic of this type of data collections is the sparseness and dynamic nature. A crucial computational task is then the ''link prediction problem'' which allows to suggest possible collaborators for scientists in social networks, or to discover unknown interactions between proteins to explain the mechanism of a disease in biological networks, or to suggest novel products to be bought to a customer in a e-commerce recommendation system. Most approaches however do not effectively exploit the topological contextual information available in the neighborhood of each edge. Here we cast the problem as a binary classification task over specific subgraphs and employ efficient graph kernels to solve it. We show empirically state-of-the-art results on several benchmark datasets.

@@

Unsupervised learning methods are non-adaptive, and hence computationally efficient, and define a score for any node pair that is proportional to the existence likelihood of an edge between the two nodes. \textit{Adamic-Adar} \cite{adamic} computes the weighted sum over the common neighbors where the weight is inversely proportional to the (log of) each neighbor node degree. The \textit{preferential attachment} method computes a score simply as the product of the node degrees in an attempt to exploit the 'rich get richer' property of certain network dynamics. \textit{Katz} \cite{katz} takes into account the number of common paths with different lengths between two nodes, assigning more weight to shorter paths. The \textit{Leicht-Holme-Newman} method \cite{lhni} computes the number of intermediate nodes. In \cite{matrix-factorization} the score is derived from the singular value decomposition of the adjacency matrix.

@@

In the link prediction problem we are given a graph $G(V,E)$ and a binary target vector $Y=\{y_{(0,0)},y_{(0,1)}, \cdots, y_{(|V|,|V|)}\}$ where $y_{(u,v)}=1$ if $(u,v) \in E$ and 0 otherwise. The training data is obtained considering a random subset of edges in $E^{tr} \in E$ and inducing a training graph $G^{tr}=(V,E^{tr})$. The remaining edges $E^{ts} = E \setminus E^{tr}$ are used to partition the target vectors: $Y^{tr} = \{y_{(u,v)} | (u,v) \in E^{tr}\}$, $Y^{ts} = \{y_{(u,v)} | (u,v) \in E^{ts}\}$.
We can now cast the problem as a standard classification problem in the domain of graphs. Given $G^{tr}$ we build a train and test set as the corresponding joint neighborhood subgraphs as detailed in Section~\ref{sec:link}. 
We can now compute a Gram matrix of the instances and solve the classification task using for example the efficient LinearSVC library \cite{svm}.

\section{Experiment}
To compare the performance of the JNSL method with other link prediction approaches we follow \cite{matrix-factorization} and use 6 datasets belonging to different domains.
\begin{itemize}
\item \textit{Protein }\cite{protein-protein}: nodes are proteins and edges encodes a thresholded interaction confidence between proteins. It has 2617 nodes and 11855 links with an average degree of 9.1.

@@

\textbf{Model Selection}: The values of different hyper-parameters are set by using a 3-fold on training set in which we use one fold for training and the rest two folds for validation. We tune the values of radius for extracting subgraphs $R$ in $\lbrace 1, 2 \rbrace$, $\lambda$ in node label function in $\lbrace 10, 15 \rbrace$, for $r$ and $d$ parameters of NSPDK in $\lbrace  1, 2 \rbrace$ and $\lbrace  1, 2, 3 \rbrace$, respectively. Finally, the regularization tradeoff $C$ for the SVM is picked up in $\lbrace 10^{-4}, 10^{-3}, 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2, 10^3,\ 10^4 \rbrace$.

@@

\section{Conclusion and Future Work}
We have presented a novel approach to link prediction in absence of side information that can effectively exploit the topological contextual information available in a neighborhood of each edge. We have empirically shown that this approach achieves very competitive results compared to other state-of-the-art methods.
In the future, we will investigate how to make use of multiple and heterogeneous information sources when these are available for nodes and edges.

>>>>

\begin{abstract}
A crucial computational task for relational and network data is the ''link prediction problem'' which allows for example to discover unknown interactions between proteins to explain the mechanism of a disease in biological networks, or to suggest novel products for a customer in a e-commerce recommendation system. Most approaches however do not effectively exploit the contextual information available in the neighborhood of each edge. Here we propose to cast the problem as a binary classification task over the union of the pair of subgraphs located at the endpoints of each edge. We model the classification task using a support vector machine endowed with an efficient graph kernel and achieve state-of-the-art results on several benchmark datasets.

@@

Unsupervised methods are non-adaptive (i.e. they do not have parameters that are tuned on the specific problem instance), and can therefore be computationally efficient. In general they define a score for any node pair that is proportional to the existence likelihood of an edge between the two nodes. \textit{Adamic-Adar} \cite{adamic} computes the weighted sum over the common neighbors where the weight is inversely proportional to the (log of) each neighbor node degree. The \textit{preferential attachment} method computes a score simply as the product of the node degrees in an attempt to exploit the 'rich get richer' property of certain network dynamics. \textit{Katz} \cite{katz} takes into account the number of common paths with different lengths between two nodes, assigning more weight to shorter paths. The \textit{Leicht-Holme-Newman} method \cite{lhni} computes the number of intermediate nodes. In \cite{matrix-factorization} the score is derived from the singular value decomposition of the adjacency matrix.

@@

In the link prediction problem we are given a graph $G(V,E)$ and a binary target vector $Y=\{y_{(0,0)},y_{(0,1)}, \cdots, y_{(|V|,|V|)}\}$ where $y_{(u,v)}=1$ if $(u,v) \in E$ and 0 otherwise. The training data is obtained considering a random subset of edges in $E^{tr} \in E$ and inducing a training graph $G^{tr}=(V,E^{tr})$. Note that the graph used for training does not contain any of the edges that will be queried in the test phase. The remaining edges $E^{ts} = E \setminus E^{tr}$ are used to partition the target vectors: $Y^{tr} = \{y_{(u,v)} | (u,v) \in E^{tr}\}$, $Y^{ts} = \{y_{(u,v)} | (u,v) \in E^{ts}\}$.
We can now cast the problem as a standard classification problem in the domain of graphs. Given $G^{tr}$ we build a train and test set as the corresponding joint neighborhood subgraphs as detailed in Section~\ref{sec:link}. 
We can now compute a Gram matrix of the instances and solve the classification task using for example the efficient LinearSVC library \cite{svm}.

\section{Empirical evaluation}
To compare the performance of the JNSL method with other link prediction approaches we follow \cite{matrix-factorization} and use 6 datasets belonging to different domains.
\begin{itemize}
\item \textit{Protein }\cite{protein-protein}: nodes are proteins and edges encodes a thresholded interaction confidence between proteins. It has 2617 nodes and 11855 links with an average degree of 9.1.

@@

\textbf{Model Selection}: The values of different hyper-parameters are set by using a 3-fold on the training set, that is, always considering only the training network, we use one fold for fitting the parameters and the rest two folds for validating the effect of the hyper parameter choice. We tune the values of radius for extracting subgraphs $R$ in $\lbrace 1, 2 \rbrace$, $\lambda$ in node label function in $\lbrace 10, 15 \rbrace$, for $r$ and $d$ parameters of NSPDK in $\lbrace  1, 2 \rbrace$ and $\lbrace  1, 2, 3 \rbrace$, respectively. Finally, the regularization tradeoff $C$ for the SVM is picked up in $\lbrace 10^{-4}, 10^{-3}, 10^{-2},\ 10^{-1}, 1,\ 10,\ 10^2, 10^3,\ 10^4 \rbrace$.

@@

\section{Conclusion and Future Work}
We have presented a novel approach to link prediction in absence of side information that can effectively exploit the topological contextual information available in the neighborhood of each edge. We have empirically shown that this approach achieves very competitive results compared to other state-of-the-art methods.
In future work, we will investigate how to make use of multiple and heterogeneous information sources when these are available for nodes and edges.
